Enforced Invariant Confluence

### 2020-01-13 doc/the_vldb_journal_paper/sections/introduction.tex
Minor introduction formatting changes.
No content changes, just some formatting changes.

\section{Introduction}
When an application designer decides to replicate a piece of data, they have to
make a fundamental choice between weak and strong consistency. Replicating the
data with strong consistency using a technique like distributed transactions
(e.g.,~\cite{bernstein1981concurrency,mohan1986transaction}) or state machine
replication (e.g.,~\cite{schneider1990implementing, lamport1998part,
liskov2012viewstamped, ongaro2014search}) makes the application designer's life
very easy. To the developer, a strongly consistent system behaves exactly like
a single-threaded system running on a single node, so reasoning about the
behavior of the system is simple~\cite{herlihy1990linearizability}.
Unfortunately, strong consistency is at odds with performance. The CAP theorem
and PACELC theorem tell us that strongly consistent systems suffer from higher
latency at best and unavailability at worst~\cite{gilbert2002brewer,
brewer2012cap, abadi2012consistency}. On the other hand, weak consistency
models like eventual consistency~\cite{vogels2009eventually}, PRAM
consistency~\cite{lipton1988pram}, causal consistency~\cite{ahamad1995causal},
and others~\cite{lloyd2011don, mehdi2017can} allow data to be replicated with
high availability and low latency, but they put a tremendous burden on the
application designer to reason about the complex interleavings of operations
that are allowed by these weak consistency models. In particular, weak
consistency models strip an application developer of one of the earliest and
most effective tools that is used to reason about the execution of programs:
application invariants~\cite{hoare1969axiomatic, balegas2015towards} such as
database integrity constraints~\cite{grefen1993integrity, gupta1993local}. Even
if every transaction executing in a weakly consistent system individually
maintains an application invariant, the system as a whole can produce
invariant-violating states.

Is it possible for us to have our strongly consistent cake and eat it with high
availability too? Can we replicate a piece of data with weak consistency but
still ensure that its invariants are maintained? Yes... sometimes. Bailis et
al.\ introduced the notion of \emph{\invariantconfluence{}} as a necessary and
sufficient condition for when invariants can be maintained over replicated data
without the need for any coordination~\cite{bailis2014coordination}. If an
object is \invariantconfluent{} with respect to an invariant, we can replicate
it with the performance benefits of weak consistency and (some of) the
correctness benefits of strong consistency.

@@

\textbf{The second main thrust of this paper is to partially avoid coordination
even in programs that require it}, by generalizing \invariantconfluence{} to a
property called \emph{segmented \invariantconfluence{}}. While
\invariantconfluence{} characterizes objects that can be replicated
\emph{without any} coordination, segmented \invariantconfluence{} allows us to
replicate non-\invariantconfluent{} objects with only \emph{occasional}
coordination. The main idea is to divide the set of invariant-satisfying states
into \emph{segments}, with a restricted set of transactions allowed in each
segment.  Within a segment servers act without any coordination; they
synchronize only to transition across segment boundaries. This design
highlights the trade-off between application complexity and
coordination-freedom; more complex applications require more segments which
require more coordination, and vice-versa.

Finally, we present Lucy: an implementation of our decision procedures and a
system for replicating \invariantconfluent{} and segmented
\invariantconfluent{} objects. Using Lucy, we find that our decision
procedures can efficiently handle a wide range of common workloads. For
example, in \secref{Evaluation}, we apply Lucy to foreign key constraints,
escrow transactions, an auction application, and the TPC-C benchmark. Lucy
processes these workloads in less than half a second, and no workload requires
more than 66 lines of code to specify. Moreover, we find that segmented
\invariantconfluent{} replication can achieve 10x to 100x more throughput than
linearizable replication for low-coordination workloads.

In closing, here is an outline of the paper and of our contributions:
%
We propose a novel expression-oriented definition of \invariantconfluence{}
that is both formal and simple (\secref{InvariantConfluence}).
%
We develop an understanding of why \invariantclosure{} is not necessary for
\invariantconfluence{} and use this understanding to develop conditions under
which it is both necessary and sufficient (\secref{InvariantClosure}).
%
We exploit these conditions to design a general-purpose interactive decision
procedure for \invariantconfluence{} (\secref{InteractiveDecisionProcedure}).
%
We again exploit these conditions to design a novel non-trivial sufficient
condition for \invariantconfluence{}, called \mergereducibility.
%
We present segmented \invariantconfluence{}: a generalization of
\invariantconfluence{} that uses a small amount of coordination to maintain
invariants for replicated objects that are otherwise not \invariantconfluent{}
(\secref{SegmentedInvariantConfluence}).
%
We evaluate our methods using a prototype implementation called Lucy
(\secref{Evaluation}).

>>>>

\section{Introduction}
When an application designer decides to replicate a piece of data, they have to
make a fundamental choice between weak and strong consistency. Replicating the
data with strong consistency using a technique like distributed transactions
(e.g.,~\cite{bernstein1981concurrency,mohan1986transaction}) or state machine
replication (e.g.,~\cite{schneider1990implementing, lamport1998part,
liskov2012viewstamped, ongaro2014search}) makes the application designer's life
very easy. To the developer, a strongly consistent system behaves exactly like
a single-threaded system running on a single node, so reasoning about the
behavior of the system is simple~\cite{herlihy1990linearizability}.
Unfortunately, strong consistency is at odds with performance. The CAP theorem
and PACELC theorem tell us that strongly consistent systems suffer from higher
latency at best and unavailability at worst~\cite{gilbert2002brewer,
brewer2012cap, abadi2012consistency}. On the other hand, weak consistency
models like eventual consistency \cite{vogels2009eventually}, PRAM
consistency~\cite{lipton1988pram}, causal consistency~\cite{ahamad1995causal},
and others~\cite{lloyd2011don, mehdi2017can} allow data to be replicated with
high availability and low latency, but they put a tremendous burden on the
application designer to reason about the complex interleavings of operations
that are allowed by these weak consistency models. In particular, weak
consistency models strip an application developer of one of the earliest and
most effective tools that is used to reason about the execution of programs:
application invariants~\cite{hoare1969axiomatic, balegas2015towards} such as
database integrity constraints~\cite{grefen1993integrity, gupta1993local}. Even
if every transaction executing in a weakly consistent system individually
maintains an application invariant, the system as a whole can produce
invariant-violating states.

Is it possible for us to have our strongly consistent cake and eat it with high
availability too? Can we replicate a piece of data with weak consistency but
still ensure that its invariants are maintained? Yes...\ sometimes. Bailis et
al.\ introduced the notion of \emph{\invariantconfluence{}} as a necessary and
sufficient condition for when invariants can be maintained over replicated data
without the need for any coordination~\cite{bailis2014coordination}. If an
object is \invariantconfluent{} with respect to an invariant, we can replicate
it with the performance benefits of weak consistency and (some of) the
correctness benefits of strong consistency.

@@

\textbf{The second main thrust of this paper is to partially avoid coordination
even in programs that require it}, by generalizing \invariantconfluence{} to a
property called \emph{segmented \invariantconfluence{}}. While
\invariantconfluence{} characterizes objects that can be replicated
\emph{without any} coordination, segmented \invariantconfluence{} allows us to
replicate non-\invariantconfluent{} objects with only \emph{occasional}
coordination. The main idea is to divide the set of invariant-satisfying states
into \emph{segments}, with a restricted set of transactions allowed in each
segment. Within a segment, servers act without any coordination; they
synchronize only to transition across segment boundaries. This design
highlights the trade-off between application complexity and
coordination-freedom; more complex applications require more segments which
require more coordination, and vice-versa.

Finally, we present Lucy: an implementation of our decision procedures and a
system for replicating \invariantconfluent{} and segmented
\invariantconfluent{} objects. Using Lucy, we find that our decision
procedures can efficiently handle a wide range of common workloads. For
example, in \secref{Evaluation}, we apply Lucy to foreign key constraints,
escrow transactions, an auction application, and the TPC-C benchmark. Lucy
processes these workloads in less than half a second, and no workload requires
more than 66 lines of code to specify. Moreover, we find that segmented
\invariantconfluent{} replication can achieve 10$\times$ to 100$\times$ more
throughput than linearizable replication for low-coordination workloads.

In closing, here is a summary of our contributions:
\begin{itemize}
  \item
    We propose a novel expression-oriented definition of \invariantconfluence{}
    that is both formal and simple (\secref{InvariantConfluence}).
  \item
    We develop an understanding of why \invariantclosure{} is not necessary for
    \invariantconfluence{} and use this understanding to develop conditions
    under which it is both necessary and sufficient
    (\secref{InvariantClosure}).
  \item
    We exploit these conditions to design a general-purpose interactive
    decision procedure for \invariantconfluence{}
    (\secref{InteractiveDecisionProcedure}).
  \item
    We again exploit these conditions to design a novel non-trivial sufficient
    condition for \invariantconfluence{}, called \mergereducibility{}
    (\secref{MergeReduction}).
  \item
    We present segmented \invariantconfluence{}: a generalization of
    \invariantconfluence{} that uses a small amount of coordination to maintain
    invariants for replicated objects that are otherwise not
    \invariantconfluent{} (\secref{SegmentedInvariantConfluence}).

  % \item
  %   We evaluate our methods using a prototype implementation called Lucy
  %   (\secref{Evaluation}).
\end{itemize}
